---
title: "Lab 1"
author: "Renee Albrecht"
date: "January 9, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)

```

read in noaa_fisheries.csv data 

```{r}
us_landings <- read_csv(here("data", "noaa_fisheries.csv"))
```


Now, make a tidy version of the dataframe 

```{r}
landings_tidy <- us_landings %>%
  clean_names() %>%
  mutate(state = str_to_lower(state),
         afs_name = str_to_lower(afs_name)) %>%
  mutate(dollars_num = parse_number(dollars_usd))

```

Let's just get some information for salmon 

```{r}
salmon_landings <- landings_tidy %>%
  mutate(afs_clean = str_remove(afs_name, pattern = "aggregate")) %>%
  filter(str_detect(afs_clean, pattern = "salmon")) %>%
  separate(afs_clean, into = c("group","species"), sep = ",")
  
```

In excel, the following is equivalent to pivot tables 

Find some grouped summary data: 

Find annual total US landings and dollar value (summing across all states) for each TYPE of salmon using 'group_by' + 'summarize()'

```{r}
salmon_summary <- salmon_landings %>%
  group_by(year, species) %>%
  summarize(
    tot_landings = sum(landings_pounds), 
    tot_value = sum(dollars_num)
  )

```

make a graph! 
```{r}
salmon_landings_graph <- ggplot(data = salmon_summary, aes(x = year, y = tot_landings, group = species)) + 
  geom_line(aes(color = species))+
  theme_minimal()

salmon_landings_graph

```

Now we will export our graph with ggsave
```{r}

ggsave(plot = salmon_landings_graph, 
       here("figures", "us_salmon_ra.png"),
       height = 5, 
       width = 8)

```

Make a nice kable table
```{r}
salmon_first_5 <- salmon_summary %>%
  head(5)

#Use kable

kable(salmon_first_5) %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

  
```

